---
title: "Simulation of ARR in Pregnancy, based on PRIMS data"
author: "Marianne Tokic"
date: "3/31/2021"
output: html_document
---





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

library(dplyr)
library(ggplot2)
library(reda)
```

## PRIMS Data

```{r}
### Data from the original PRIMS Paper
dat.PRIMS <- data.frame(
  timeframe = c(             "trim1.yp" 
                            ,"trim2.yp"   
                            ,"trim3.yp" 
                            ,"trim4.yp" 
                            ,"trim1.preg" 
                            ,"trim2.preg" 
                            ,"trim3.preg"
                           , "trim1.pp"   
                           , "trim2.pp"   
                            ,"trim3.pp"   
                           , "trim4.pp")
  
  ,tf = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11)
  , ARR = c(
    0.6
    ,0.7
    ,0.7
    ,0.8
    
    ,0.5
    ,0.6
    ,0.2
    
    ,1.2
    ,0.9
    ,0.9
    ,0.6
  )
  , lower.CI = c(
    0.5
    ,0.6
    ,0.6
    ,0.7
    
    ,0.4
    ,0.5
    ,0.2
    
    ,1.1
    ,0.8
    ,0.8
    ,0.5
  )
  ,upper.CI = c(
    0.7
    ,0.8
    ,0.9
    ,0.9
    
    ,0.6
    ,0.7
    ,0.3
    
    ,1.4
    ,1.0
    ,1.0
    ,0.7
  )
  ,n = rep(227, 11)
  , dur.tf = c(rep(91.25, 4), 84, 112, 84,rep(91.25, 4) ) # add duration for each timeframe
) %>% 

## compute SD from CI, assuming normal distribution (which is NOT true)
  mutate(SD= sqrt(n) * (upper.CI - lower.CI)/3.92)

dat.PRIMS <- dat.PRIMS %>% 
  mutate(ARR_simple= c(0.7, 0.7, 0.7, 0.7, 0.5, 0.6, 0.2, 1.2, 0.9, 0.7, 0.7))
```

```{r}

ggplot2::ggplot( data=dat.PRIMS, aes(x=tf, y=ARR)) + 
  geom_rect( mapping=aes(xmin=4.8, xmax=7.2, ymin=0, ymax=0.75, fill= "pregnancy"), color="grey", alpha=0.5)+
  geom_line()+
  geom_errorbar(aes(ymin=lower.CI, ymax=upper.CI), width=.1) +
  ggtitle("PRIMS data, 95%CI as reported")
  
```

In the PRIMS study Vrusic et al. (2004) described the disease activity in N= 227 (mostly) untreated RR-MS patients during pregnancy. The above graph shows their findings as described by the mean annualized relapse rate (ARR) in their cohort.

This study was the first to demonstrate that while we ob observe a post-partum peak in disease activity, the ARR drops drastically in the pregnancy - mostpronounced in the third trimster.

This is what has become known as the "classic course of disease" in pregnancy in context of RR-MS.

However, since then 
-the diagnostic standarts have changed
-therapies have become more effective
-therapies are introduced earlier
-- this might lead to an overall more beging disease
-non DMT  treatment of RR-MS patients has advanced

And so, it is unclear wetherr this "classic" course of disease is still representative of modern days RR-MS in context of pregnancy.

## Goal 1: Simulating the baseline hazard as presented in PRIMS
In this doctument, I try to generate a simulated data set based on the PRIMS data. In the long run I want to supplement more modern data sourced fromt he KuMS registry. However, untilt he data in the registry is cleaned up and tested fo plausibility, I supplement it with the PRIMS data.

### Requirement for the data
-the course of the ARR in pregnancy should be simulated as a base riks rate in different timeframes
-relapses (=events) should occure atleast 30 days apart -> Interarrivalt time (IAT)
-population mean ARR per timeframe should roughly correspond to the PRIMS data
-the format of te data should be in counting process format AND in survival format to be able to later compare diffent analysis methods

## simulation 1: baseline hazard via means of the PRIMS study, no smoothing, no IAT modification, no covariats


```{r}

## Set simulation parameters
N_IDs = 100
N_loops = 10
 
## Set times
trimester1 = 84
trimester2= 112
trimester3= 84
year = 365
quarter_year = year/4
pregnancy = trimester1 + trimester2 + trimester3

duration = year + pregnancy + year

```

Simulation of `r N_IDs` x `r N_loops ` pregnancies. baseline hazard sourced via the mean ARRS in the PRIMS study.
```{r}
####bazeline hazard = rhoFun, full time simulated

# 
# 
# ## Set simulation parameters
# N_IDs = 100
# N_loops = 10
# 
# ## Set functions for simulation
# # baseline hazard
# rhoFun = function(timevector, b) { 
#   timepoint <- case_when( 
#     
#     timevector <= year 
#     ~ 4, 
#     
#     timevector > year &
#     timevector <= year + trimester1
#     ~ 5, 
#     
#     timevector > year + trimester1 &
#     timevector <= year + trimester1 + trimester2
#     ~ 6, 
#     
#     timevector > year + trimester1 + trimester2 &
#     timevector <= year + pregnancy
#     ~ 7,
#     
#     timevector > year + pregnancy &
#     timevector <= year + pregnancy + quarter_year
#     ~ 8,
#     
#     timevector > year + pregnancy + quarter_year &
#     timevector <= year + pregnancy + quarter_year + quarter_year
#     ~ 9,
#     
#     timevector > year + pregnancy + quarter_year + quarter_year
#     ~ 10,
#   )
#   return( b[timepoint]/year)
#   
# }
# 
# ## make dataframe to save simulated data
# dat.sim <- data.frame(ID = NA
#                       ,time = NA
#                       , event = NA
#                       , origin = NA
#                       , X = NA
#                       )
# 
# ## Simulation
# for (counter in 1:N_loops) {
# sim <- simEventData(N_IDs
#              ,method = "inversion"
#              ,origin= 1
#              ,endTime = duration
#              ,rho = rhoFun
#              ,arguments = list(
#                rho = list(b = dat.PRIMS$ARR_simple)
#                )
#   )
# 
# sim <- sim %>% 
#   mutate(ID = ID + (counter*N_loops*N_IDs))
# 
# 
# dat.sim <-  rbind(dat.sim, sim)
# }
# 
# dat.sim <- dat.sim %>% # remove the NA entry
#   filter(!is.na(ID))
# 
# save(dat.sim, file = "dat.sim")

load("dat.sim")
```

```{r}
## sorting events to timeframes

IDs <- dat.sim %>% # get all IDs
  select(ID) %>% unique()

tmp <- dat.sim %>% # get all events
  filter(event == 1) %>% 
  mutate(
    tf = case_when( # sort evets into timeframes
      
      time <= quarter_year
      ~ 1,
      
      time <= quarter_year*2 &
      time > quarter_year
      ~ 2,
      
      
    time <= quarter_year*3 &
    time > quarter_year*2
      ~ 3,
    
    time > quarter_year*3 &
    time <= year
    ~ 4,
    
    time > year &
    time <= year + trimester1
    ~ 5,
    
    time > year + trimester1 &
    time <= year+ trimester1 + trimester2
    ~ 6,
    
    time > year + trimester1+ trimester2 &
    time <= year + pregnancy
    ~ 7,
    
    time > year + pregnancy &
    time <= year + pregnancy + quarter_year
    ~ 8,
    
    time > year + pregnancy + quarter_year &
    time <= year + pregnancy + (2*quarter_year)
    ~ 9,
    
    time > year + pregnancy +(2*quarter_year) &
    time <= year + pregnancy + (3*quarter_year)
    ~ 10,
    
    time > year + pregnancy +(3*quarter_year) &
    time <= year + pregnancy + year
    ~ 11,
  )
)

num_events <- tmp %>% 
  group_by(tf) %>% 
  add_count(ID, name = "num_events") %>% 
  ungroup() %>% 
  select(ID, tf, num_events) %>% unique()

dat.sim_counting <- data.frame(ID = NA
                      ,tf = NA)


for (counter in 1:11) {
  
  tmp2 <- IDs %>% 
    mutate(
      tf = counter
    )
  
  
 dat.sim_counting <- rbind(dat.sim_counting, tmp2) 
}

dat.sim_counting <- dat.sim_counting %>% # remove the NA entry
  filter(!is.na(ID)) %>% 
  arrange(ID)

dat.sim_counting <- left_join( dat.sim_counting, num_events) %>% 
  arrange(ID) %>% 
  mutate(num_events = ifelse(is.na(num_events), 0, num_events)) %>% 
  mutate(ditf = case_when(
    
    tf == 5
    ~ trimester1,
    
    tf == 6
    ~ trimester2,
    
    tf == 7
    ~ trimester3,
    
    T ~ quarter_year
  ))
```

```{r}
# get mean ARR per tf
tmp <- dat.sim_counting %>% 
  mutate(ARR = 
           ifelse(num_events > 0, num_events/ditf * year, 0)
         ) %>% 
  select(ID, ARR, tf)

tmp2 <- tidyr::pivot_wider(tmp
                    , id_cols = ID
                    , names_from = tf
                    , values_from = ARR
                    )

tmp3 <- data.frame(
  tf = 1:11
  ,ARR = c(
    mean(tmp2$`1`)
    ,mean(tmp2$`2`)
    ,mean(tmp2$`3`)
    ,mean(tmp2$`4`)
    ,mean(tmp2$`5`)
    ,mean(tmp2$`6`)
    ,mean(tmp2$`7`)
    ,mean(tmp2$`8`)
    ,mean(tmp2$`9`)
    ,mean(tmp2$`10`)
    ,mean(tmp2$`11`)
  )
  , lower.CL  = c(
    t.test(tmp2$`1`)$"conf.int"[1]
    , t.test(tmp2$`2`)$"conf.int"[1]
    , t.test(tmp2$`3`)$"conf.int"[1]
    , t.test(tmp2$`4`)$"conf.int"[1]
    , t.test(tmp2$`5`)$"conf.int"[1]
    , t.test(tmp2$`6`)$"conf.int"[1]
    , t.test(tmp2$`7`)$"conf.int"[1]
    , t.test(tmp2$`8`)$"conf.int"[1]
    , t.test(tmp2$`9`)$"conf.int"[1]
    , t.test(tmp2$`10`)$"conf.int"[1]
    , t.test(tmp2$`11`)$"conf.int"[1]
  )
  , upper.CL  = c(
    t.test(tmp2$`1`)$"conf.int"[2]
    , t.test(tmp2$`2`)$"conf.int"[2]
    , t.test(tmp2$`3`)$"conf.int"[2]
    , t.test(tmp2$`4`)$"conf.int"[2]
    , t.test(tmp2$`5`)$"conf.int"[2]
    , t.test(tmp2$`6`)$"conf.int"[2]
    , t.test(tmp2$`7`)$"conf.int"[2]
    , t.test(tmp2$`8`)$"conf.int"[2]
    , t.test(tmp2$`9`)$"conf.int"[2]
    , t.test(tmp2$`10`)$"conf.int"[2]
    , t.test(tmp2$`11`)$"conf.int"[2]
  )
)

```

```{r}
#plot
ggplot2::ggplot( data=tmp3, aes(x=tf, y=ARR)) + 
  geom_rect( mapping=aes(xmin=4.8, xmax=7.2, ymin=0, ymax=0.75, fill= "pregnancy"), color="grey", alpha=0.5)+
  geom_line(color = "blue")+
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=.1, color ="blue") +
  ggtitle("Simulated data, 95%CI for the mean, no IAT mod")
  
  
```

This looks very good. 
Here is the number IDs per number of events
```{r}
tmp <- dat.sim %>% 
  group_by(ID) %>% 
  mutate( num_events = sum(event)
         )

hist(tmp$num_events)

```
```{r}
m.sim1 <- lme4::glmer(num_events ~ as.factor(tf) 
                      + (1|ID) -1
            , offset = log(ditf/year)
            , data = dat.sim_counting
            , family = poisson()
            )

sjPlot::tab_model(m.sim1)
```


```{r}
# with(dat.sim_counting,table(as.factor(tf),num_events))
# 
# library(dotwhisker)
# library(broom.mixed)
# dwplot(m.sim1,effects="fixed")
# library(ggplot2)
# theme_set(theme_bw())
# ggplot(dat.sim_counting,aes(num_events,tf,colour=num_events))+
#     stat_sum(alpha=0.5)
```

## simulation 2: baseline hazard via means of the PRIMS study, no smoothing,  IAT modification, no covariats

```{r}
 ## Set simulation parameters
 N_IDs = 10
 N_loops = 100
 
 ## Set times
trimester1 = 84
trimester2= 112
trimester3= 84
year = 365
quarter_year = year/4
pregnancy = trimester1 + trimester2 + trimester3

duration = year + pregnancy + year
```


Simulation of `r N_IDs` x `r N_loops ` pregnancies. baseline hazard sourced via the mean ARRS in the PRIMS study.
```{r}
####bazeline hazard = rhoFun, full time simulated, IAT simulation



# ## Set functions for simulation
# # baseline hazard
#  rhoFun = function(timevector, b) { 
#    timepoint <- case_when( 
#      
#     timevector <= year 
#      ~ 4, 
#      
#      timevector > year &
#      timevector <= year + trimester1
#      ~ 5, 
#      
#      timevector > year + trimester1 &
#      timevector <= year + trimester1 + trimester2
#      ~ 6, 
#     
#      timevector > year + trimester1 + trimester2 &
#      timevector <= year + pregnancy
#      ~ 7,
#     
#      timevector > year + pregnancy &
#     timevector <= year + pregnancy + quarter_year
#      ~ 8,
#     
#     timevector > year + pregnancy + quarter_year &
#     timevector <= year + pregnancy + quarter_year + quarter_year
#      ~ 9,
#      
#     timevector > year + pregnancy + quarter_year + quarter_year
#      ~ 10,
#    )
#    return( b[timepoint]/year)
#   
#  }
#  ## IAT function
#  custom_IA <- function(rate) {
#   truncnorm::rtruncnorm(n=1, a=30, b=Inf, mean=400, sd=13)
# }
#  
#  ## make dataframe to save simulated data
# dat.sim2 <- data.frame(ID = NA
#                        ,time = NA
#                        , event = NA
#                        , origin = NA
#                        , X = NA
#                        )
#  
#  ## Simulation
#  for (counter in 1:N_loops) {
#  sim <- simEventData(N_IDs
#               ,method = "inversion"
#               ,origin= 1
#               ,endTime = duration
#               ,rho = rhoFun
#               ,interarrival = custom_IA
#               ,arguments = list(
#                rho = list(b = dat.PRIMS$ARR_simple)
#                 )
#    )
#  
#  sim <- sim %>% 
#    mutate(ID = ID + (counter*N_loops*N_IDs))
#  
# 
#  dat.sim2 <-  rbind(dat.sim2, sim)
#  }
#  
#  dat.sim2 <- dat.sim2 %>% # remove the NA entry
#    filter(!is.na(ID))
#  
#  save(dat.sim2, file = "dat.sim2")

 load("dat.sim2")
```


```{r}
## sorting events to timeframes

IDs <- dat.sim2 %>% # get all IDs
  select(ID) %>% unique()

tmp <- dat.sim2 %>% # get all events
  filter(event == 1) %>% 
  mutate(
    tf = case_when( # sort evets into timeframes
      
      time <= quarter_year
      ~ 1,
      
      time <= quarter_year*2 &
      time > quarter_year
      ~ 2,
      
      
    time <= quarter_year*3 &
    time > quarter_year*2
      ~ 3,
    
    time > quarter_year*3 &
    time <= year
    ~ 4,
    
    time > year &
    time <= year + trimester1
    ~ 5,
    
    time > year + trimester1 &
    time <= year+ trimester1 + trimester2
    ~ 6,
    
    time > year + trimester1+ trimester2 &
    time <= year + pregnancy
    ~ 7,
    
    time > year + pregnancy &
    time <= year + pregnancy + quarter_year
    ~ 8,
    
    time > year + pregnancy + quarter_year &
    time <= year + pregnancy + (2*quarter_year)
    ~ 9,
    
    time > year + pregnancy +(2*quarter_year) &
    time <= year + pregnancy + (3*quarter_year)
    ~ 10,
    
    time > year + pregnancy +(3*quarter_year) &
    time <= year + pregnancy + year
    ~ 11,
  )
)

num_events <- tmp %>% 
  group_by(tf) %>% 
  add_count(ID, name = "num_events") %>% 
  ungroup() %>% 
  select(ID, tf, num_events) %>% unique()

dat.sim_counting <- data.frame(ID = NA
                      ,tf = NA)


for (counter in 1:11) {
  
  tmp2 <- IDs %>% 
    mutate(
      tf = counter
    )
  
  
 dat.sim_counting <- rbind(dat.sim_counting, tmp2) 
}

dat.sim_counting <- dat.sim_counting %>% # remove the NA entry
  filter(!is.na(ID)) %>% 
  arrange(ID)

dat.sim_counting <- left_join( dat.sim_counting, num_events) %>% 
  arrange(ID) %>% 
  mutate(num_events = ifelse(is.na(num_events), 0, num_events)) %>% 
  mutate(ditf = case_when(
    
    tf == 5
    ~ trimester1,
    
    tf == 6
    ~ trimester2,
    
    tf == 7
    ~ trimester3,
    
    T ~ quarter_year
  ))
```

```{r}
# get mean ARR per tf
tmp <- dat.sim_counting %>% 
  mutate(ARR = 
           ifelse(num_events > 0, num_events/ditf * year, 0)
         ) %>% 
  select(ID, ARR, tf)

tmp2 <- tidyr::pivot_wider(tmp
                    , id_cols = ID
                    , names_from = tf
                    , values_from = ARR
                    )

tmp3 <- data.frame(
  tf = 1:11
  ,ARR = c(
    mean(tmp2$`1`)
    ,mean(tmp2$`2`)
    ,mean(tmp2$`3`)
    ,mean(tmp2$`4`)
    ,mean(tmp2$`5`)
    ,mean(tmp2$`6`)
    ,mean(tmp2$`7`)
    ,mean(tmp2$`8`)
    ,mean(tmp2$`9`)
    ,mean(tmp2$`10`)
    ,mean(tmp2$`11`)
  )
  , lower.CL  = c(
    t.test(tmp2$`1`)$"conf.int"[1]
    , t.test(tmp2$`2`)$"conf.int"[1]
    , t.test(tmp2$`3`)$"conf.int"[1]
    , t.test(tmp2$`4`)$"conf.int"[1]
    , t.test(tmp2$`5`)$"conf.int"[1]
    , t.test(tmp2$`6`)$"conf.int"[1]
    , t.test(tmp2$`7`)$"conf.int"[1]
    , t.test(tmp2$`8`)$"conf.int"[1]
    , t.test(tmp2$`9`)$"conf.int"[1]
    , t.test(tmp2$`10`)$"conf.int"[1]
    , t.test(tmp2$`11`)$"conf.int"[1]
  )
  , upper.CL  = c(
    t.test(tmp2$`1`)$"conf.int"[2]
    , t.test(tmp2$`2`)$"conf.int"[2]
    , t.test(tmp2$`3`)$"conf.int"[2]
    , t.test(tmp2$`4`)$"conf.int"[2]
    , t.test(tmp2$`5`)$"conf.int"[2]
    , t.test(tmp2$`6`)$"conf.int"[2]
    , t.test(tmp2$`7`)$"conf.int"[2]
    , t.test(tmp2$`8`)$"conf.int"[2]
    , t.test(tmp2$`9`)$"conf.int"[2]
    , t.test(tmp2$`10`)$"conf.int"[2]
    , t.test(tmp2$`11`)$"conf.int"[2]
  )
)

```

```{r}
#plot
ggplot2::ggplot( data=tmp3, aes(x=tf, y=ARR)) + 
  geom_rect( mapping=aes(xmin=4.8, xmax=7.2, ymin=0, ymax=0.75, fill= "pregnancy"), color="grey", alpha=0.5)+
  geom_line(color = "blue")+
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=.1, color ="blue") +
  ggtitle("Simulated data, 95%CI for the mean, IAT rtruncnorm(mean 400, sd 13)")
  
  
```

This looks about right. However, it's strange, that every ID has exactly 2 events:
```{r}
tmp <- dat.sim2 %>% 
  group_by(ID) %>% 
  mutate( num_events = sum(event)
         )

hist(tmp$num_events)


```


## Poisson of the simulated data

```{r}
m.sim1 <- lme4::glmer(num_events ~ as.factor(tf) 
                      + (tf|ID) -1
            , offset = log(ditf/year)
            , data = dat.sim_counting
            , family = poisson()
            )
```

```{r}
lme4::isSingular(m.sim1, tol = 1e-4)
```

From lme4:
"omplex mixed-effect models (i.e., those with a large number of variance-covariance parameters) frequently result in singular fits, i.e. estimated variance-covariance matrices with less than full rank. Less technically, this means that some "dimensions" of the variance-covariance matrix have been estimated as exactly zero. For scalar random effects such as intercept-only models, or 2-dimensional random effects such as intercept+slope models, singularity is relatively easy to detect because it leads to random-effect variance estimates of (nearly) zero, or estimates of correlations that are (almost) exactly -1 or 1. However, for more complex models (variance-covariance matrices of dimension >=3) singularity can be hard to detect; models can often be singular without any of their individual variances being close to zero or correlations being close to +/-1.

While singular models are statistically well defined (it is theoretically sensible for the true maximum likelihood estimate to correspond to a singular fit), there are real concerns that (1) singular fits correspond to overfitted models that may have poor power; (2) chances of numerical problems and mis-convergence are higher for singular models (e.g. it may be computationally difficult to compute profile confidence intervals for such models); (3) standard inferential procedures such as Wald statistics and likelihood ratio tests may be inappropriate."

So let's look at the model...

```{r}
sjPlot::tab_model(m.sim1)
```

As expected, I find exactly the mean ARR we observed, which are nearly identical to effects put into the simulation:
```{r}
tmp <- tmp3 %>% 
  left_join(dat.PRIMS %>% select(tf, ARR_simple))

tmp2 <- summary(m.sim1)$coefficients %>% as.data.frame()

tmp <- tmp %>% 
  cbind(., exp(tmp2$Estimate))

names(tmp) <- c(
  "time frame"
  , "observed ARR"
  , "observed lower CL"
  , "observed uppler CL"
  , "baseline hazard"
  , "estimated ARR"
)


kableExtra::kable(tmp)
```


## survival analyisis: Andersen-Gills

```{r}

## to do : coding of the data set into survival data with timedependent covariat: tf
# this data formating SUCKS so much!


```

```{r}
# need to ne redone
# library(survival)
# surv1 <- with(dat.sim_surv, Surv(time, event))
# 
# m.sim2 <- coxph(surv1 ~ as.factor(tf), data = dat.sim_surv)
# 
# summary(m.sim2)

```

