---
title: "First Test"
author: "Marianne Tokic"
date: "19 3 2021"
output: html_document
---

```{r setup, include=FALSE}
# IMPORTANT! THIS IS A PACKRAT PROJECT!!! #
### otions
knitr::opts_chunk$set(echo = F)

### packages
library(tidyr)
library(dplyr)
library(simrec)
library(varhandle)



```
## simrec

The function simrec allows simulation of recurrent event data following the multiplicative intensity model described in Andersen and Gill (Andersen and Gill 1982) with the baseline hazard being a function of the total/calendar time. To induce between-subject-heterogeneity a random effect covariate (frailty term) can be incorporated. Data for individual i are generated according to the intensity process
Yi(t)⋅λ0(t)⋅Zi⋅exp(βtXi)
where Xi defines the covariate vector, and β the regression coefficient vector. λ0(t) denotes the baseline hazard, being a function of the total/calendar time t, and Yi(t) the predictable process that equals one as long as individual i is under observation and at risk for experiencing events. Zi denotes the frailty variable with (Zi)i iid with E(Zi)=1 and Var(Zi)=θ. The parameter θ describes the degree of between-subject-heterogeneity. Data output is in the counting process format.
from vingiette

### Simulation of a stable ARR

Here I try to simulate just a stable ARR across a observational period. This means the baseline hazard doesn't change over time and no Covariats are included.
Simulation of just relpase data: about a risk of 0.5 per year across group, min. 30 days apart

```{r}
#### parameters

### observation parameters

N_goal <-  100000 #numers of subjects wanted
N <- 100 # Number of simulated Subjects per run off the loop
N_loops <- N_goal/N
##
year <-  365
pregnancy <- 280
years <- round((2*year + pregnancy) / year, digits = 2) # set observation time in years
ARR <- 0.5

## Between-Subject differences
# The parameter θ describes the degree of between-subject-heterogeneity
# a gamma distributed frailty variable with variance 0.25
dist.z <- "gamma"
#par.z <-  0.25 #like described above
par.z <- 0 # turned off

## baseline hazard
## here I try to just hit a ARR of 0.5
# 1. Weibull-shaped baseline hazard with shape parameter lambda=1
# and scale parameter nu=2.
#dist.rec <- "weibull"
#par.rec <- c(0.5,2)

#2. step baseline hazard
# with lambda= a = 0.5until t1 and lambda = b = 0.5 after t1
# t1 is set to one day, so the mean ARR *should* be 0.5
dist.rec="step"
a= 0.5 / year
b= 0.5 / year
t_1 = 1
par.rec=c(a,b,t_1)

## Follow up time
# Subjects are to be followed for two years with 20\% of the subjects
# being censored according to a uniformly distributed censoring time
# within [0,2] (in years).

#fu.min <- years
#fu.max <- years
#cens.prob <- 0.2

#no censoring, follow up for 2 years plus pregnancy
fu.min <- 2*year + pregnancy
fu.max <- 2*year + pregnancy
cens.prob <- 0

## Risk-free period
# After each event a subject is not at risk for experiencing further events
## for a period of 30 days 

dfree <- 30
pfree <- 1

## Covariats, set to have no influence = beta == 0
dist.x <- "binomial"
par.x  <- 0.5
beta.x <-  0
```

```{r}
#simulation loop
for (count_loops in 1:N_loops) {
  if(count_loops == 1)
  {simdata <- simrec(N, fu.min, fu.max, cens.prob, dist.x , par.x , beta.x,
                  dist.z, par.z, dist.rec, par.rec, pfree, dfree)
  
  simdata$id <- simdata$id + N*count_loops
  }
  tmp <- simrec(N, fu.min, fu.max, cens.prob, dist.x , par.x , beta.x,
                  dist.z, par.z, dist.rec, par.rec, pfree, dfree)
  
  tmp$id <- tmp$id + N*count_loops
  
  simdata <- rbind(simdata, tmp)
  
  rm(tmp)
  
}
nrow(simdata %>% select(id) %>% unique()) # first check

```


```{r}
#### Simulation of a single set of N Subjects
# simdata <- simrec(N, fu.min, fu.max, cens.prob, dist.x , par.x , beta.x,
#                   dist.z, par.z, dist.rec, par.rec, pfree, dfree)
# 
# print(simdata[1:10,])
```



```{r}
# check wether the target mean ARR has been hit
simdata1 <- simdata %>% 
  group_by(id) %>% 
  mutate(num_events = sum(status)) %>% # get the number of events per ID
  ungroup() %>% 
  mutate(sim.ARR = num_events/fu * year)

tmp <- simdata1 %>% 
  dplyr::select(id, sim.ARR) %>% 
  unique()

mean(tmp$sim.ARR)
```

```{r}
### reformat data to the format as I find it in the KuMS data
simdata2 <-  simdata1 %>% 
  mutate(
         LMP = year + 1, # set trimesters
         DOD = LMP + pregnancy,
         
         end_trim1.year_prior = year/4,
         end_trim2.year_prior = (year/4) * 2,
         end_trim3.year_prior = (year/4) * 3,
         
         end_trim1.preg = LMP + 84,
         end_trim2.preg = end_trim1.preg + 112,
         
         end_trim1.post_partum = DOD + year/4,
         end_trim2.post_partum = DOD + (year/4) * 2,
         end_trim3.post_partum = DOD + (year/4) * 3) %>% 
         
         mutate(timeframe = case_when(
           
           stop <= end_trim1.year_prior
            ~ "trim1.yp",
           
           stop > end_trim1.year_prior &
           stop <= end_trim2.year_prior 
            ~ "trim2.yp",
           
           stop > end_trim2.year_prior &
           stop <= end_trim3.year_prior 
            ~ "trim3.yp",
           
           stop > end_trim3.year_prior &
           stop <= LMP 
            ~ "trim4.yp",
           
           stop > LMP &
           stop <= end_trim1.preg 
            ~ "trim1.preg",
           
           stop > end_trim1.preg &
           stop <= end_trim2.preg 
            ~ "trim2.preg",
           
           stop > end_trim2.preg &
           stop <= DOD 
            ~ "trim3.preg",
           
           stop > DOD &
           stop <= end_trim1.post_partum 
            ~ "trim1.pp",
           
           stop > end_trim1.post_partum &
           stop <= end_trim2.post_partum 
            ~ "trim2.pp",
           
           stop > end_trim2.post_partum &
           stop <= end_trim3.post_partum 
            ~ "trim3.pp",
           
           stop > end_trim3.post_partum 
            ~ "trim4.pp"
         ))
         



print(simdata2[1:10,])
```

```{r}
### get the counts per timeframe
 tmp <- simdata2 %>% 
  filter(status == 1) %>%  #remove all times without an event
  group_by(id) %>%
  add_count(timeframe) %>% 
  select(-c(start, stop, status, num_events)) %>% 
  unique()


```

```{r}
#get the base data for all subjects
base <-  simdata2 %>% 
  select(id
         , x
         , z
         , sim.ARR
         , LMP
         , DOD
         
         ) %>% unique()
```

```{r}
tmp2 <- base %>% 
  mutate(timeframe = "trim1.yp")

tmp3 <- left_join(base, tmp2)

tmp2 <- base %>% 
  mutate(timeframe = "trim2.yp")

tmp3 <- rbind(tmp3, tmp2)

tmp2 <- base %>% 
  mutate(timeframe = "trim3.yp")

tmp3 <- rbind(tmp3, tmp2)

tmp2 <- base %>% 
  mutate(timeframe = "trim4.yp")

tmp3 <- rbind(tmp3, tmp2)

tmp2 <- base %>% 
  mutate(timeframe = "trim1.preg")

tmp3 <- rbind(tmp3, tmp2)

tmp2 <- base %>% 
  mutate(timeframe = "trim2.preg")

tmp3 <- rbind(tmp3, tmp2)

tmp2 <- base %>% 
  mutate(timeframe = "trim3.preg")

tmp3 <- rbind(tmp3, tmp2)

tmp2 <- base %>% 
  mutate(timeframe = "trim1.pp")

tmp3 <- rbind(tmp3, tmp2)

tmp2 <- base %>% 
  mutate(timeframe = "trim2.pp")

tmp3 <- rbind(tmp3, tmp2)
tmp2 <- base %>% 
  mutate(timeframe = "trim3.pp")

tmp3 <- rbind(tmp3, tmp2)
tmp2 <- base %>% 
  mutate(timeframe = "trim4.pp")

tmp3 <- rbind(tmp3, tmp2)

### make time frame a factor and set the correct order
tmp3$timeframe <- as.factor(tmp3$timeframe)

tmp3$timeframe  <- factor(tmp3$timeframe 
                ,levels = c( "trim1.yp" 
                            ,"trim2.yp"   
                            ,"trim3.yp" 
                            ,"trim4.yp" 
                            ,"trim1.preg" 
                            ,"trim2.preg" 
                            ,"trim3.preg"
                           , "trim1.pp"   
                           , "trim2.pp"   
                            ,"trim3.pp"   
                           , "trim4.pp" ))
```

```{r}
### make final, full dataframe
simpois <-  tmp3 %>% 
  left_join(., simdata2 %>% select(-c(start, stop, status, num_events, timeframe)) %>% unique() ) %>% 
  left_join(., tmp) %>% 
  mutate(events_in_timeframe = ifelse(is.na(n), 0, n)
         ) %>% 
  mutate(days_in_timeframe = case_when(
    timeframe == "trim1.yp"
    ~ year/4,
    
    timeframe == "trim2.yp"
    ~ year/4,
    
    timeframe == "trim3.yp"
    ~ year/4,
    
    timeframe == "trim4.yp"
    ~ year/4,
    
    timeframe == "trim1.preg"
    ~ 84,
    
    timeframe == "trim2.preg"
    ~ 112,
    
    timeframe == "trim3.preg"
    ~ 84,
    
    timeframe == "trim1.pp"
    ~ year/4,
    
    timeframe == "trim2.pp"
    ~ year/4,
    
    timeframe == "trim3.pp"
    ~ year/4,
    
    timeframe == "trim4.pp"
    ~ year/4,
  ))
```

```{r}
### cleanup
rm.all.but(c("simdata2", "simpois"))
```
# plot events per day to look for patterns

```{r}
tmp  <- simdata2   %>% 
  filter(status == 1) %>% 
  select(stop) %>% unique()

tmp$day <- round(tmp$stop) 

# 
# tmp <- transform(tmp, freq.day = ave(seq(nrow(tmp)), day, FUN=length)) %>% 
#   select(-c(id, stop)) %>% unique()
```

```{r}
hist(x = tmp$day, freq = T, breaks = 11)
```



# looking at the plots like in the KuMS Project to compare

might give a signularity warning, but that's just what we want - all timeframes should be the same. Thus, we also do not expect any significant  betas

```{r}
simpois$timeframe <- as.factor(simpois$timeframe)

simpois$timeframe  <- factor(simpois$timeframe 
                ,levels = c( "trim1.yp" 
                            ,"trim2.yp"   
                            ,"trim3.yp" 
                            ,"trim4.yp" 
                            ,"trim1.preg" 
                            ,"trim2.preg" 
                            ,"trim3.preg"
                           , "trim1.pp"   
                           , "trim2.pp"   
                            ,"trim3.pp"   
                           , "trim4.pp" ))

simpois$id <- as.character(simpois$id)


m1 <- lme4::glmer(
  events_in_timeframe ~ timeframe  + (1|id) + offset(log(days_in_timeframe/365))
  , data =  simpois
  , family = poisson()
)
```

```{r}
summary(m1)
```


```{r}
### make time frame a factor and set the correct order



emmeans::emmip(m1,   ~ timeframe, type = "response", offset=0 , CIs = T, data = simpois)
```


