---
title: "Simulation of pregnancy relapse rates"
author: "Marianne Tokic"
date: "3/20/2021"
output: html_document
---

```{r setup, include=FALSE}
# IMPORTANT! THIS IS A PACKRAT PROJECT!!! #
### options
knitr::opts_chunk$set(echo = F)

### packages
library(tidyr)
library(dplyr)
library(simrec)
library(varhandle)
library(ggplot2)
library(Runuran)

```
## Overview of this skript
In this script I simply try to simulate the natural course of the disease through pregnancy as per the PRIMS data.
In the first step, I generate a big data set of mean ARRs in the range of the values found in the PRIMS study.
Then,  I try to get a good description of the course in the PRIMS data by:
- cubic linear regression
- natuarl splines
- b-splines
- ...?

In the last step, I try to integrate the description of the disease course into either simrec or simEvent

## Natural course of the disease as per PRIMS:
getting a bigger sample of ARRs like in the PRIMS data

```{r}
### Data from the original PRIMS Paper
dat.PRIMS <- data.frame(
  timeframe = c(             "trim1.yp" 
                            ,"trim2.yp"   
                            ,"trim3.yp" 
                            ,"trim4.yp" 
                            ,"trim1.preg" 
                            ,"trim2.preg" 
                            ,"trim3.preg"
                           , "trim1.pp"   
                           , "trim2.pp"   
                            ,"trim3.pp"   
                           , "trim4.pp")
  
  ,tf = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11)
  , ARR = c(
    0.6
    ,0.7
    ,0.7
    ,0.8
    
    ,0.5
    ,0.6
    ,0.2
    
    ,1.2
    ,0.9
    ,0.9
    ,0.6
  )
  , lower.CI = c(
    0.5
    ,0.6
    ,0.6
    ,0.7
    
    ,0.4
    ,0.5
    ,0.2
    
    ,1.1
    ,0.8
    ,0.8
    ,0.5
  )
  ,upper.CI = c(
    0.7
    ,0.8
    ,0.9
    ,0.9
    
    ,0.6
    ,0.7
    ,0.3
    
    ,1.4
    ,1.0
    ,1.0
    ,0.7
  )
  ,n = rep(227, 11)
  , dur.tf = c(rep(91.25, 4), 84, 112, 84,rep(91.25, 4) ) # add duration for each timeframe
) %>% 

## compute SD from CI, assuming normal distribution (which is NOT true)
  mutate(SD= sqrt(n) * (upper.CI - lower.CI)/3.92)

dat.PRIMS <- dat.PRIMS %>% 
  mutate(ARR_simple= c(0.7, 0.7, 0.7, 0.7, 0.5, 0.6, 0.2, 1.2, 0.9, 0.7, 0.7))
```

```{r}
ggplot2::qplot(tf, ARR, data=dat.PRIMS, geom="line") + 
    geom_errorbar(aes(ymin=lower.CI, ymax=upper.CI), width=.1)
```

### generate data points that roughly fit the PRIMS data 

I assume poisson distribution at this point. Also, for all timeframes ther minimun number of events is 0 and the maximum number of relapses is 3 for the "long" periods and 2 for the first and last pregnancy trimester



```{r}
### simulate data like found in the PRIMS data, here only one data set is produced
N = 1000
sim.PRIMS <- data.frame(ID = 1:N)

for (counter in 1:nrow(dat.PRIMS)) {
  if(dat.PRIMS$dur.tf[counter] >= 90){ # long time frames = 3 events possible
    
    #dat.PRIMS$ARR[counter] * dat.PRIMS$dur.tf[counter]) / 365 to sale back from ARR to rate!
    
    dist <- udpois((dat.PRIMS$ARR_simple[counter] * dat.PRIMS$dur.tf[counter]) / 365 , 0, 3)
    gen <- darid.new(dist)
    tmp <- as.data.frame(ur(gen, N))
    names(tmp) <- paste0("count")
    
    tmp <- tmp%>% 
      mutate(ID = row_number()) %>% 
      mutate(ditf = dat.PRIMS$dur.tf[counter])
    
    tmp <- tmp %>% 
      mutate(ARR =(count / ditf )* 365 )
       
    names(tmp) <- c(paste0("event_count_",counter), "ID", paste0("ditf_",counter), paste0("ARR_",counter))
    sim.PRIMS <-  left_join(sim.PRIMS, tmp)
  }
  if(dat.PRIMS$dur.tf[counter] <90){ # short timeframes, 2 events possible
    
    dist <- udpois((dat.PRIMS$ARR_simple[counter] * dat.PRIMS$dur.tf[counter]) / 365, 0, 2)
    gen <- darid.new(dist)
    tmp <- as.data.frame(ur(gen, N))
    names(tmp) <- paste0("count")
    
    tmp <- tmp%>% 
      mutate(ID = row_number()) %>% 
      mutate(ditf = dat.PRIMS$dur.tf[counter])
    
    tmp <- tmp %>% 
      mutate(ARR =(count / ditf )* 365 )
       
    names(tmp) <- c(paste0("event_count_",counter), "ID", paste0("ditf_",counter), paste0("ARR_",counter))
    sim.PRIMS <-  left_join(sim.PRIMS, tmp)
  }
}

sum(sim.PRIMS$ARR_5) / N


```

### simulate multiple ARRs pre timeframe like in the PRIMS data
```{r}
# simulate multiple data sets to get multiple mean ARRs per timeframe

N_datasets = 1000
N_IDs = 1000

sim.ARR.PRIMS <- data.frame(ARR_1 = NA,    
                            ARR_2 = NA,    
                            ARR_3 = NA,     
                            ARR_4  = NA,    
                            ARR_5  = NA,   
                            ARR_6  = NA,   
                            ARR_7 = NA,     
                            ARR_8 = NA,   
                            ARR_9 = NA,  
                            ARR_10 = NA,   
                            ARR_11  = NA)

for (count_sets in 1:N_datasets) {
  sim.PRIMS <- data.frame(ID = 1:N_IDs)
  
  for (counter in 1:nrow(dat.PRIMS)) {
  if(dat.PRIMS$dur.tf[counter] >= 90){ # long time frames = 3 events possible
    
    #dat.PRIMS$ARR[counter] * dat.PRIMS$dur.tf[counter]) / 365 to sale back from ARR to rate!
    
    dist <- udpois((dat.PRIMS$ARR_simple[counter] * dat.PRIMS$dur.tf[counter]) / 365 , 0, 3)
    gen <- darid.new(dist)
    tmp <- as.data.frame(ur(gen, N_IDs))
    names(tmp) <- paste0("count")
    
    tmp <- tmp%>% 
      mutate(ID = row_number()) %>% 
      mutate(ditf = dat.PRIMS$dur.tf[counter])
    
    tmp <- tmp %>% 
      mutate(ARR =(count / ditf )* 365 )
       
    names(tmp) <- c(paste0("event_count_",counter), "ID", paste0("ditf_",counter), paste0("ARR_",counter))
    sim.PRIMS <-  left_join(sim.PRIMS, tmp)
  }
  if(dat.PRIMS$dur.tf[counter] <90){ # short timeframes, 2 events possible
    
    dist <- udpois((dat.PRIMS$ARR_simple[counter] * dat.PRIMS$dur.tf[counter]) / 365, 0, 2)
    gen <- darid.new(dist)
    tmp <- as.data.frame(ur(gen, N_IDs))
    names(tmp) <- paste0("count")
    
    tmp <- tmp%>% 
      mutate(ID = row_number()) %>% 
      mutate(ditf = dat.PRIMS$dur.tf[counter])
    
    tmp <- tmp %>% 
      mutate(ARR =(count / ditf )* 365 )
       
    names(tmp) <- c(paste0("event_count_",counter), "ID", paste0("ditf_",counter), paste0("ARR_",counter))
    sim.PRIMS <-  left_join(sim.PRIMS, tmp)
  }
  tmp <- sim.PRIMS %>% 
    select(contains("ARR"))
  
  tmp <- colMeans(tmp)
  }
sim.ARR.PRIMS <-  rbind(sim.ARR.PRIMS, tmp) 
}

sim.ARR.PRIMS <- sim.ARR.PRIMS %>% 
  filter(!is.na(ARR_1)) 

```


```{r}
sim.ARR.PRIMS_long <- pivot_longer(
  sim.ARR.PRIMS
  ,everything()
  ,names_to = "tf"
  ,values_to = "simulated_mean_ARR"
) %>% 
  mutate(tf = readr::parse_number(tf))
```


```{r}
qplot(tf, simulated_mean_ARR, data=sim.ARR.PRIMS_long, geom="point") #simulated PRIMS ARR in black
last_plot() + geom_line(aes(x=tf, y=ARR), data = dat.PRIMS, col=5) # original PRIMS data in light blue
```

looks right---

## Fitting different curves to the data

### cubic linear model
```{r}
# I think one possible specification would be a cubic linear model
y.hat <- predict(lm(simulated_mean_ARR~tf+I(tf^2)+I(tf^3), data=sim.ARR.PRIMS_long))
qplot(tf, simulated_mean_ARR, data=sim.ARR.PRIMS_long, geom="point") #simulated PRIMS ARR in black
last_plot() + geom_line(aes(x=tf, y=y.hat), col=2) # the fitted values red lines
last_plot() + geom_line(aes(x=tf, y=ARR), data = dat.PRIMS, col=5) # original PRIMS data in light blue
```

Not really a good summary of the  data....

### 5th order polynomial

```{r}
# or try a 5th order polynomial
y.hat <- predict(lm(simulated_mean_ARR ~ poly(tf, 5, raw=TRUE), data=sim.ARR.PRIMS_long)) # estimating the model and obtaining the fitted values from the model

qplot(tf, simulated_mean_ARR, data=sim.ARR.PRIMS_long, geom="point") #simulated PRIMS ARR in black
last_plot() + geom_line(aes(x=tf, y=y.hat), col=2) # the fitted values red lines
last_plot() + geom_line(aes(x=tf, y=ARR), data = dat.PRIMS, col=5) # original PRIMS data in light blue

```


Looks better, but not good. 

### Splines:



```{r}
#loading the Splines Packages
require(splines)

attach(dat.PRIMS) #attaching  dataset

tflims<-range(tf)
tf.grid<-seq(from=tflims[1], to = tflims[2])


#cutpoints at each tf arounf the pregnancy 4, 5, 6, 7
fit<-lm(ARR ~ bs(tf,knots = c(4, 5, 6, 7)),data = dat.PRIMS )
        
summary(fit)

#Plotting the Regression Line to the scatterplot   
plot(tf,ARR,col="grey",xlab="tf",ylab="ARR")
points(tf.grid,predict(fit,newdata = list(tf=tf.grid)),col="darkgreen",lwd=2,type="l")
#adding cutpoints
abline(v=c(25,40,60),lty=2,col="darkgreen")
```

```{r}
#fitting smoothing splines using smooth.spline(X,Y,df=...)
fit1<-smooth.spline(tf,ARR,df=10) #6 degrees of freedom
#Plotting both cubic and Smoothing Splines 
plot(tf,ARR,col="grey",xlab="tf",ylab="ARR")
points(tf.grid,predict(fit,newdata = list(tf=tf.grid)),col="darkgreen",lwd=2,type="l")
#adding cutpoints

lines(fit1,col="red",lwd=2)
legend("topright",c("Smoothing Spline with 10 df","Cubic Spline"),col=c("red","darkgreen"),lwd=2)
```

The smoothing spline looks best up to now. Maybe try on a simplyfied data set with stable ARR pre pregnancy and simpler drop post partum ?

### smoothing splines with 10 df on "simple" ARR


```{r}
#fitting smoothing splines using smooth.spline(X,Y,df=...)
fit1<-smooth.spline(tf,ARR_simple,df=10) #6 degrees of freedom
#Plotting both cubic and Smoothing Splines 
plot(tf,ARR_simple,col="grey",xlab="tf",ylab="ARR")

lines(fit1,col="red",lwd=2)
legend("topright",c("Smoothing Spline with 10 df"),col=c("red"),lwd=2)

detach(dat.PRIMS)
```
 Perfect :) :) :) This, I can use !

## base chosen method on long genrated dataset
```{r}
sim.ARR.PRIMS_long <- sim.ARR.PRIMS_long %>% 
  arrange(tf) %>% 
  mutate(index = row_number())

```

```{r}
s

# or try a b-spline
y.hat <- predict(lm(simulated_mean_ARR ~ bs(index, 8), data=sim.ARR.PRIMS_long)) # estimating the model and obtaining the fitted values from the model

qplot(tf, ARR_simple, data=dat.PRIMS, geom="point") #simulated PRIMS ARR in black
last_plot() + geom_line(aes(x=index/1000, y=y.hat), data = sim.ARR.PRIMS_long , col=2) # the fitted values red lines
last_plot() + geom_line(aes(x=tf, y=ARR), data = dat.PRIMS, col=5) # original PRIMS data in light blue

```

```{r}
attach(sim.ARR.PRIMS_long)

#fitting smoothing splines using smooth.spline(X,Y,df=...)
fit1<-smooth.spline(index,simulated_mean_ARR,df=10) #6 degrees of freedom
#Plotting both cubic and Smoothing Splines 
detach(sim.ARR.PRIMS_long)
qplot(tf, ARR_simple, data=dat.PRIMS, geom="point") #simulated PRIMS ARR in black
last_plot() + geom_line(aes(x=index/1000, y=fit1$y), data = sim.ARR.PRIMS_long , col=2) # the fitted values red lines
last_plot() + geom_line(aes(x=tf, y=ARR), data = dat.PRIMS, col=5) # original PRIMS data in light blue
```

## simEvent test

```{r}
base.hazard <- dat.PRIMS %>% 
  select(tf, dur.tf, ARR_simple) %>% 
  mutate(rate = ARR_simple/365)
```


```{r}
### get an mean ARR of 0.7 across 100 subjects
library(reda)
test <- simEventData(100
            ,method = "inversion"
             ,origin= 1
             ,endTime = 366
             ,rho = 0.7 /365
             )
```

```{r}

test <- simEventData(10
            ,method = "inversion"
             ,origin= 1
             ,endTime = 11
             ,rho = function(timeVec) {
               tmp <- as.data.frame(trunc(timeVec)) %>% mutate(y = NA);
               for (i in 1:nrow(tmp)) {
                 tmp$y[i] <- case_when(
                   
                   tmp[i,1] == 1
                   ~ base.hazard$rate[1],
                   tmp[i,1] == 2
                   ~ base.hazard$rate[2],
                   tmp[i,1] == 3
                   ~ base.hazard$rate[3],
                   tmp[i,1] == 4
                   ~ base.hazard$rate[4],
                   tmp[i,1] == 5
                   ~ base.hazard$rate[5],
                   tmp[i,1] == 6
                   ~ base.hazard$rate[6],
                   tmp[i,1] == 7
                   ~ base.hazard$rate[7],
                   tmp[i,1] == 8
                   ~ base.hazard$rate[8],
                   tmp[i,1] == 9
                   ~ base.hazard$rate[9],
                   tmp[i,1] == 10
                   ~ base.hazard$rate[10],
                   tmp[i,1] == 11
                   ~ base.hazard$rate[11],
                   tmp[i,1] == 12
                   ~ base.hazard$rate[12],
                   
                 )
               }
              ; return(tmp$y)
             }
             )
```


```{r}
test2 <- test %>% 
  filter(event == 1) %>% 
  mutate(tf = trunc(time)) %>% 
  group_by(tf) %>% 
  add_count(tf) %>% 
  arrange(tf) %>% 
  mutate(ARR = ((n/10/90))*365)

tmp <- test2 %>% 
  select(tf, ARR) %>% 
  unique()
```


