---
title: "Sim_test_3"
author: "Marianne Tokic"
date: "3/21/2021"
output: html_document
---


```{r setup, include=FALSE}
# IMPORTANT! THIS IS A PACKRAT PROJECT!!! #
### options
knitr::opts_chunk$set(echo = F)

### packages
library(tidyr)
library(dplyr)
library(simrec)
library(varhandle)
library(ggplot2)
library(Runuran)
library(reda)
```

```{r}
N= 100
pregnancy = 280
year = 365
```

## Overview

in this script, I test out the simEvent function in the reda package, as simrec() doesn't allow for user specified baseline hazard rates.
However, in simEvent() the rho-function is poorly defined and described. As such, I want to test out, whether it's feasable to (a) set the baseline hazard so, that there's an ARR of 0.7 as base and then (b) to model the pregnancy influence as a covariate

## baseline hazard that leads to an ARR = 0.7
### ARR = 0.7 in one year
```{r}
### get an mean ARR of 0.7 across 100 subjects

test <- simEventData(1000
            ,method = "inversion"
             ,origin= 1
             ,endTime = 366
             ,rho = 0.7 /365
             )
tmp <- test %>% 
  select(ID) %>% unique()

tmp2 <- test %>% 
  filter(event == 1) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), 365, time)
         ) %>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()

mean(tmp3$num_events)
```

this works reliably well.

### ARR = 0.7 in 2 years plus pregnancy

```{r}
### get an mean ARR of 0.7 across 100 subjects

test <- simEventData(N
            ,method = "inversion"
             ,origin= 1
             ,endTime = pregnancy + 2*year
             ,rho = 0.7 /365 #baseline hazard, scaled to DAY
             )
tmp <- test %>% 
  select(ID) %>% unique()

tmp2 <- test %>% 
  filter(event == 1) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), 365, time)
         ) %>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()

mean(tmp3$num_events)/ (pregnancy + 2*year) * year
```

this works, too. It's improtant to keep the time in this function as DAYS and also scale the hazard to days.
So, ARR to harzard per day = ARR/365!

## Modulate the hazard in pregnancy

### stable reduction in pregnancy (= days 366 - 646) to an ARR of 0.2

```{r}
### get an ARR of 0.7 in year pre pregnancy and post-partum. Reduce ARR in pregnancy to 0.2

## set parameters for covariat = pregnancy effect
### time-varying covariates and time-varying coefficients
zFun <- function(time) {
 as.numeric(time > 366 & time <= 646 ) #0 = outside of pregnancy; 1= in pregnancy
}




test <- simEventData(100
             ,method = "inversion"
             ,origin= 1
             ,endTime = pregnancy + 2*year
             ,rho =  0.7/year  #baseline hazard or event rate, scaled to DAY, ARR_roh/year
             
             ,z =  zFun
             ,zCoef =  -0.5#(-0.5 * pregnancy)/year # works like this:
                          # ARR_rho +  zCoef
                          #for example: ARR_rho = 0.5, zCoef = 0.5  -> ARR in pregnancy = 1
             
            # , interarrival = function(n, rate) rgamma(n=1, shape =  rate*1000)
             )

## get total ARR
tmp <- test %>% 
  select(ID) %>% unique()

tmp2 <- test %>% 
  filter(event == 1) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), 365, time)
         ) %>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()

print(
  paste0("ARR total= ", mean((tmp3$num_events/(pregnancy + 2*year))*year)
  )
)

## get ARR year pre pregnancy
tmp <- test %>% 
  select(ID) %>% unique()

tmp2 <- test %>% 
  filter(event == 1 & time < 366) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), 365, time)
         )%>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()

print(
  paste0("ARR prior to pregnancy= ", mean(tmp3$num_events)
  )
)

## get ARR in pregnancy
tmp <- test %>% 
  select(ID) %>% unique()

tmp2 <- test %>% 
  filter(event == 1 & time > 366 & time <= 646 ) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), 365, time)
         )%>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()

print(
  paste0("ARR in pregnancy= ", mean((tmp3$num_events/pregnancy)*year)
  )
)

## get ARR post-partum
tmp <- test %>% 
  select(ID) %>% unique()

tmp2 <- test %>% 
  filter(event == 1 & time > 646) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), 365, time)
         )%>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()

print(
  paste0("ARR post-partum= ", mean((tmp3$num_events/year)*year)
  )
)

```

Now, I let this run in a loop to get 1000 estimated of the ARRs to inspect the simulation even more

```{r include=FALSE}
N_loops = 100
N_IDs = 100

base.haz = 1
preg.effect = log(0.6)
duration = pregnancy + 2*year

Simulated.ARRs = data.frame(
  Set = rep(NA, N_loops)
  , ARR.pre = rep(NA, N_loops)
  , ARR.in = rep(NA, N_loops)
  ,ARR.post = rep(NA, N_loops)
  ,ARR.total = rep(NA, N_loops)
)

zFun <- function(time) {
 as.numeric(time > 366 & time <= 646 ) #0 = outside of pregnancy; 1= in pregnancy
}


for (counter in 1:N_loops) {
## save number of the loop
  Simulated.ARRs$Set[counter] <-  counter
  

## simulate the data
sim <- simEventData(N_IDs
             ,method = "inversion"
             ,origin= 1
             ,endTime = duration
             ,rho =  base.haz/year  #baseline hazard or event rate, scaled to DAY, ARR_roh/year
             
             ,z =  zFun
             ,zCoef =  preg.effect
             )


### get the ARRs

## get total ARR
tmp <- sim %>% 
  select(ID) %>% unique()

tmp2 <- sim %>% 
  filter(event == 1) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), 365, time)
         ) %>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()
#save
Simulated.ARRs$ARR.total[counter]=  mean((tmp3$num_events/(pregnancy + 2*year))*year)



## get ARR year pre pregnancy
tmp <- sim %>% 
  select(ID) %>% unique()

tmp2 <- sim %>% 
  filter(event == 1 & time < 366) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), 365, time)
         )%>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()

#save
Simulated.ARRs$ARR.pre[counter]=  mean((tmp3$num_events/year)*year)



## get ARR in pregnancy
tmp <- sim %>% 
  select(ID) %>% unique()

tmp2 <- sim %>% 
  filter(event == 1 & time > 366 & time <= 646 ) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), 365, time)
         )%>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()
#save
Simulated.ARRs$ARR.in[counter]=  mean((tmp3$num_events/pregnancy)*year)



## get ARR post-partum
tmp <- sim %>% 
  select(ID) %>% unique()

tmp2 <- sim %>% 
  filter(event == 1 & time > 646) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), 365, time)
         )%>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()
#save
Simulated.ARRs$ARR.post[counter]=  mean((tmp3$num_events/year)*year)
}


```

```{r}
# checking how the zCoef interacts wth roh
#Sim1 <- Simulated.ARRs # zCoef = -0.5 -> ARR.im ca 0.425 => 0.275 difference to base.haz
mean(Simulated.ARRs$ARR.in)
mean(Simulated.ARRs$ARR.pre)
#Sim2<- Simulated.ARRs # zCoef = -0.7 -> ARR.in ca 0.343
#Sim3<- Simulated.ARRs # zCoef = -1 -> ARR.in ca 0.249=> 0.275 difference
#Sim4<- Simulated.ARRs # zCoef = -1.5 -> ARR.in ca 0.156
#Sim5<- Simulated.ARRs # baseha z = 1, zCoef = -1.5 -> ARR.in ca 0.211 , exp(-1) = 0.368
#Sim6<- Simulated.ARRs # baseha z = 1, zCoef = -2 -> ARR.in ca 0.137, exp(-2) = 0.135
#Sim7<- Simulated.ARRs # baseha z = 1, zCoef = 0.5 -> ARR.in ca 1.667
#Sim8<- Simulated.ARRs # baseha z = 1, zCoef = -0.5/year -> ARR.in ca 1...
#Sim9<- Simulated.ARRs # baseha z = 1, zCoef = -0.5*year -> ARR.in ca 0.0001 ?
Sim10<- Simulated.ARRs # baseha z = 1, zCoef = log(0.6) -> ARR.in ca 0.59
```

This interaction off effects is really puzzeling to me. Because of this I'll run a set of simpler simulations to eliminate possible sources error. Hiere I set roh = 1 and just simulate one year. Also, the Effekt of the covariat kicks in at day 1 

```{r include=FALSE}
#simpler simulation to understand the effects
N_loops = 100
N_IDs = 100

base.haz = 0.7
preg.effect = log(0.29)
duration = year 

Simulated2.ARR = data.frame(
  Set = rep(NA, N_loops)
  , ARR = rep(NA, N_loops)
)

zFun <- function(time) {
 as.numeric(time > 1 ) 
}


for (counter in 1:N_loops) {
## save number of the loop
Simulated2.ARR$Set[counter] <-  counter
  

## simulate the data
sim <- simEventData(N_IDs
             ,method = "inversion"
             ,origin= 1
             ,endTime = duration
             ,rho =  base.haz/year  #baseline hazard or event rate, scaled to DAY, ARR_roh/year
             
             ,z =  zFun
             ,zCoef =  preg.effect
             )


### get the ARR

## get total ARR
tmp <- sim %>% 
  select(ID) %>% unique()

tmp2 <- sim %>% 
  filter(event == 1) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), 365, time)
         ) %>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()
#save
Simulated2.ARR$ARR[counter]=  mean((tmp3$num_events/duration) *year)

}


```

```{r}
# checking how the zCoef interacts wth roh

mean(Simulated2.ARR$ARR)

#Sim.Simple<- Simulated.ARRs #base.haz = 1, zCoef = 0.5 -> ARR(simulated) = 1.6 
#Sim.Simple1<- Simulated.ARRs #base.haz = 1, zCoef = 0 -> ARR(simulated) = 1.0062
#Sim.Simple2<- Simulated.ARRs #base.haz = 1, zCoef = 2 -> ARR(simulated) = 7.37 ???? das entspricht base.haz * exp(zCoef), bzw 1 * 1.38
## Wie gebe ich dann die zCoefs an? als log(Zielwert) ?
# test mit der Notation log(Zielwert)
#Sim.Simple3<- Simulated.ARRs #base.haz = 1, zCoef = log(0.5) -> ARR(simulated) = 0.49
## Für positive Werte Funktioniert das sehr gut! da wir base.haz * exp(zCoef) haben, geben zCoef >0 aber <1 eine Reduktion in  ARR
#Sim.Simple4<- Simulated.ARRs #base.haz = 1, zCoef = log(0.2) -> ARR(simulated) = 0.19
# schaffe ich es bei einem base.haz von 0.7 die ARR auf 0.2 zu reduzieren? Dafür rechne ich ZielARR/base.haz = 0.2857 = 0.29
Sim.Simple5<- Simulated.ARRs #base.haz = 0.7, zCoef = log(0.29) -> ARR(simulated) = 0.2047
### YEAH!!! Das funktioniert!
```

Als nächstes möchte ich nund ie Erkentnisse daraus auf die große Simutation übertragen. Im ersten Schritt passe ich nur die ARR in der Schwangerschaft an die PRIMS Daten an mittels einer zCoef Funktion:

```{r}
### Data from the original PRIMS Paper
dat.PRIMS <- data.frame(
  timeframe = c(             "trim1.yp" 
                            ,"trim2.yp"   
                            ,"trim3.yp" 
                            ,"trim4.yp" 
                            ,"trim1.preg" 
                            ,"trim2.preg" 
                            ,"trim3.preg"
                           , "trim1.pp"   
                           , "trim2.pp"   
                            ,"trim3.pp"   
                           , "trim4.pp")
  
  ,tf = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11)
  , ARR = c(
    0.6
    ,0.7
    ,0.7
    ,0.8
    
    ,0.5
    ,0.6
    ,0.2
    
    ,1.2
    ,0.9
    ,0.9
    ,0.6
  )
  , lower.CI = c(
    0.5
    ,0.6
    ,0.6
    ,0.7
    
    ,0.4
    ,0.5
    ,0.2
    
    ,1.1
    ,0.8
    ,0.8
    ,0.5
  )
  ,upper.CI = c(
    0.7
    ,0.8
    ,0.9
    ,0.9
    
    ,0.6
    ,0.7
    ,0.3
    
    ,1.4
    ,1.0
    ,1.0
    ,0.7
  )
  ,n = rep(227, 11)
  , dur.tf = c(rep(91.25, 4), 84, 112, 84,rep(91.25, 4) ) # add duration for each timeframe
) %>% 

## compute SD from CI, assuming normal distribution (which is NOT true)
  mutate(SD= sqrt(n) * (upper.CI - lower.CI)/3.92)

dat.PRIMS <- dat.PRIMS %>% 
  mutate(ARR_simple= c(0.7, 0.7, 0.7, 0.7, 0.5, 0.6, 0.2, 1.2, 0.9, 0.7, 0.7))
```


```{r include=FALSE}
#simpler simulation to understand the effects
N_loops = 10
N_IDs = 100

trimester1 = 84
trimester2= 112
trimester3= 84

pregnancy = trimester1 + trimester2 + trimester3

base.haz = 0.7
#preg.effect = log(0.29)
duration = pregnancy + 2 * year 

Simulated2.ARR = data.frame(
  Set = rep(NA, N_loops)
  , ARR.in = rep(NA, N_loops)
  , ARR.total = rep(NA, N_loops)
)

zFun <- function(time) { # just an effect for the pregnancy, is = 1 in pregnancy and =0 in all other parts
 as.numeric(time > year & time <= pregnancy + year ) 
  #ifelse(time > 1, 1, 0)
}

zCoefFun <- function(time) { # let the effect of the Preganancy vary over time
  #for the first try, let me see how the log/exp stuff interacts with the function formualtion -> it need to be log(X)!
  # ifelse(time > year, log(0.29), 0)
  
  #now try to simulate with the PRIMS data for pregnancy
  case_when(
    time > year & 
    time <= year + trimester1
    ~ log( dat.PRIMS[5,3]/base.haz), # trimester 1
        
    time > year + trimester1 & 
    time <= year + trimester1 + trimester3
    ~ log( dat.PRIMS[6,3]/base.haz), # trimester 2
    
    time > year + trimester1 + trimester2 & 
    time <= year + pregnancy
    ~ log(dat.PRIMS[7,3]/base.haz), # trimester 3
  
  # time > year + 84 & 
  #   time <= year + 84 + 112
  #   ~ log( dat.PRIMS[5,3]/base.haz/pregnancy), #sixth timepoint, ARR cloumn in the PRIMS data
  # 
  # time > year + 84 & 
  #   time <= year + 84 + 112
  #   ~ log( dat.PRIMS[5,3]/base.haz/pregnancy), #seventh timepoint, ARR cloumn in the PRIMS data
  # 
  T ~0
  )
  }


for (counter in 1:N_loops) {
## save number of the loop
Simulated2.ARR$Set[counter] <-  counter


## simulate the data
sim <- simEventData(N_IDs
             ,method = "inversion"
             ,origin= 1
             ,endTime = duration
             ,rho =  base.haz/year  #baseline hazard or event rate, scaled to DAY, ARR_roh/year
             
             ,z =  zFun
             ,zCoef =  zCoefFun
             )


### get the ARRs

## get total ARR
tmp <- sim %>% 
  select(ID) %>% unique()

tmp2 <- sim %>% 
  filter(event == 1) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), duration, time)
         ) %>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()
#save
Simulated2.ARR$ARR.total[counter]=  mean(tmp3$num_events)/ duration *year

## get total ARR
tmp <- sim %>% 
  select(ID) %>% unique()

tmp2 <- sim %>% 
  filter(event == 1 &
        X == 1) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), duration, time)
         ) %>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()
#save
Simulated2.ARR$ARR.in[counter]=  mean(tmp3$num_events) / pregnancy * year

}
expected.rate = (dat.PRIMS[5,3] * 84 + dat.PRIMS[6,3] * 112 + dat.PRIMS[7,3] * 84)/280

mean(Simulated2.ARR$ARR.in)
expected.rate 


```

Das scheint bisher ganz gut zu funktionieren...Im nächsten Schritt möchte ich gerne versuchen, die ARR post-partum anzupassen.

## Post partum Gipfel modellieren


```{r include=FALSE}
#simpler simulation to understand the effects
N_loops = 1
N_IDs = 100

trimester1 = 84
trimester2= 112
trimester3= 84

pregnancy = trimester1 + trimester2 + trimester3

base.haz = 0.7

duration = pregnancy + 2 * year 

Simulated2.ARR = data.frame(
  Set = rep(NA, N_loops)
  , ARR.in = rep(NA, N_loops)
  , ARR.total = rep(NA, N_loops)
)

zFun <- function(time) { # effect in pregnancy and postpartum
 as.numeric(time > year #& time <= pregnancy + year 
            ) 
}

zCoefFun <- function(time) { # let the effect of the Preganancy and post partum vary over time

  case_when(
    time > year & 
    time <= year + trimester1
    ~ log( dat.PRIMS[5,3]/base.haz), # trimester 1
        
    time > year + trimester1 & 
    time <= year + trimester1 + trimester3
    ~ log( dat.PRIMS[6,3]/base.haz), # trimester 2
    
    time > year + trimester1 + trimester2 & 
    time <= year + pregnancy
    ~ log(dat.PRIMS[7,3]/base.haz), # trimester 3
    
    time > year + pregnancy & 
    time <= year + pregnancy + year/4
    ~ log(dat.PRIMS[8,3]/base.haz), # pp 1
    
    # time > year + pregnancy + year/4 & 
    # time <= year + pregnancy + year/2
    # ~ log(dat.PRIMS[9,3]/base.haz), # pp 2
    # 
    # time > year + pregnancy + year/2 & 
    # time <= year + pregnancy + (year/4) *3
    # ~ log(dat.PRIMS[10,3]/base.haz), # pp 3
    # 
    # time > year + pregnancy + (year/4) *3 & 
    # time <= duration
    # ~ log(dat.PRIMS[11,3]/base.haz), # pp 4
    
    
    T ~0
  )
}


for (counter in 1:N_loops) {
## save number of the loop
Simulated2.ARR$Set[counter] <-  counter


## simulate the data
sim <- simEventData(N_IDs
             ,method = "inversion"
             ,origin= 1
             ,endTime = duration
             ,rho =  base.haz/year  #baseline hazard or event rate, scaled to DAY, ARR_roh/year
             
             ,z =  zFun
             ,zCoef =  zCoefFun
             )


### get the ARRs

## get total ARR
tmp <- sim %>% 
  select(ID) %>% unique()

tmp2 <- sim %>% 
  filter(event == 1) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), duration, time)
         ) %>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()
#save
Simulated2.ARR$ARR.total[counter]=  mean(tmp3$num_events)/ duration *year

## get total ARR
tmp <- sim %>% 
  select(ID) %>% unique()

tmp2 <- sim %>% 
  filter(event == 1 &
        X == 1) 

tmp <- left_join(tmp, tmp2) %>% 
  select(ID, event, time) %>% 
  mutate(event= ifelse(is.na(event), 0, event),
         time= ifelse(is.na(time), duration, time)
         ) %>% 
  add_count(ID, name = "num_events") %>% 
  mutate(num_events = ifelse(event == 0, 0, num_events))

tmp3 <- tmp %>% 
  select(ID, num_events) %>% unique()
#save
Simulated2.ARR$ARR.in[counter]=  mean(tmp3$num_events) / pregnancy * year

}
expected.rate = (dat.PRIMS[5,3] * 84 + dat.PRIMS[6,3] * 112 + dat.PRIMS[7,3] * 84)/280

mean(Simulated2.ARR$ARR.in)
expected.rate 


```


## Versuch, die PRIMS-Daten als "rho" in die Simulation ein zu bauen
ich möchte das gerne noch mal probieren, ob das nicht doch funktioniert - das würde vieles einfacher machen ...

### NUR ein Jahr, noch keine anderem Zeiträume, reiner POC
```{r include=FALSE}
# bazeline hazard = rhoFun test
N_IDs = 100
N_loops = 100
trimester1 = 84
trimester2= 112
trimester3= 84

pregnancy = trimester1 + trimester2 + trimester3

duration = year 


rhoFun = function(timevector, b) { 
  timepoint <- case_when(
    timevector <= year
    ~ 5
  )
  return( b[timepoint]/year)
  
}

test <- data.frame(sim_num = rep(NA, N_loops), ARR = rep(NA, N_loops))

for (counter in 1:N_loops) {
## simulate the data
sim <- simEventData(N_IDs
             ,method = "inversion"
             ,origin= 1
             ,endTime = year
             ,rho = rhoFun
             ,arguments = list(
               rho = list(b = dat.PRIMS$ARR_simple)
               )
  )

## get ARR in pregnancy = expected 0.5
tmp <- sim %>%  # get all IDs
  select(ID) %>% unique()

tmp2 <- sim %>%  # get the number of events in pregnancy
  filter(event == 1 ) %>% 
  add_count(ID, name = "num_events") %>% 
  dplyr::select(ID, num_events) %>% 
  unique()

tmp <- left_join(tmp, tmp2) %>% 

  mutate(num_events = ifelse(is.na(num_events), 0, num_events)
         ,ARR = (num_events  / duration) * year
         )

test$sim_num[counter] <- counter
test$ARR[counter] <- mean(tmp$ARR)

}

print(mean(test$ARR))
```
Works! 
